Helmet Foam Case Study Report

### 2024-05-06 CS2 Final Draft.qmd
Edit Appendix

@@

combined
```
In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle. The dot colors in @fig-2 represent which foam sample they were measured from. As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart. So stiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.

@@

```{r}
#| echo: false
#| label: fig-3
#| warning: false
#| fig-cap: "log(stiffnessFrac) vs log(cycle) by freq, and log(dampingFrac) vs log(cycle) by freq"
#| fig-align: "center"
#| fig-width: 10
#| fig-height: 5

@@

In order to understand how log(stiffnessFrac) and log(dampingFrac) behave given different frequencies, we also colored the samples in @fig-3 by frequency. For log(stiffnessFrac), we see that higher frequencies have a more negative trend than lower frequencies. For log(dampingFrac), we see that lower frequencies have a more negative trend than higher frequencies. So frequency looks related to both stiffnessFrac and dampingFrac.

```{r}
#| echo: false
#| label: fig-4
#| warning: false
#| fig-cap: "log(stiffnessFrac) vs log(cycle) by chemIndex and porosity"
#| fig-align: "center"
#| fig-width: 18
#| fig-height: 8
#| 
sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

@@

```{r}
#| echo: false
#| label: fig-5
#| warning: false
#| fig-cap: "log(dampingFrac) vs log(cycle) for chemIndex and porosity"
#| fig-align: "center"
#| fig-width: 18
#| fig-height: 8
#| 
dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

@@

@@

### R Script
```{r ref.label=c('front-matter', 'fig-example1','tbl-summarystats', 'fig-example2')}
#| echo: true
#| eval: false
foam_data <- read.csv("foamData.csv")

foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0)

>>>>

\begin{center}
\text{Table 3: Foam Sample Type Counts}
\end{center}

@@

combined
```
In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle. The dot colors in @fig-2 represent which foam sample they were measured from. As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart. StiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.

@@

```{r}
#| echo: false
#| label: fig-3
#| warning: false
#| fig-cap: "Log(stiffnessFrac) vs Log(cycle) and Log(dampingFrac) vs Log(cycle), Colored by Freq"
#| fig-align: "center"
#| fig-width: 10
#| fig-height: 5

@@

In order to understand how log(stiffnessFrac) and log(dampingFrac) behave given different frequencies, we also colored the samples in @fig-3 by frequency. For log(stiffnessFrac), we see that higher frequencies have a more negative trend than lower frequencies. For log(dampingFrac), we see that lower frequencies have a more negative trend than higher frequencies. Frequency looks related to both stiffnessFrac and dampingFrac.

```{r}
#| echo: false
#| label: fig-4
#| warning: false
#| fig-cap: "Log(stiffnessFrac) vs Log(cycle), Separated by ChemIndex and Porosity"
#| fig-align: "center"
#| fig-width: 18
#| fig-height: 8
#| 
sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

@@

```{r}
#| echo: false
#| label: fig-5
#| warning: false
#| fig-cap: "Log(dampingFrac) vs Log(cycle), Separated by ChemIndex and Porosity"
#| fig-align: "center"
#| fig-width: 18
#| fig-height: 8
#| 
dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

@@

\begin{center}
\text{Table 4: Predictors Selected For Each Model}
\end{center}

@@

### Data Wrangling

```{r}
#| echo: true
#| eval: false
knitr::opts_chunk$set(echo = TRUE)

library(patchwork)
library("readxl")
library(ggplot2)
library(tidyverse)

foamDataRaw <- read_excel("DMA_Fatigue_Data_from_Team_Wendy_Polyurethane
                          _LLNL-MI-860456.xlsx")
names(foamDataRaw) <- foamDataRaw[3,]
foamDataRaw <- foamDataRaw[-c(1:3),] # remove rows without data
foamDataRaw <- foamDataRaw[,-c(62:64)] # remove duplicate columns

foamDataIntermediate <- foamDataRaw |>
  separate(col=SampleCode, into=c("Chemical", "chemIndex", "porosity", "ID")) |>
  select(!Chemical) |>
  mutate_at(vars(chemIndex), as.character) |>
  mutate_at(vars(porosity), as.character) |>
  mutate(chemIndex = dplyr::recode(chemIndex, "79"="1", "100"="2", "121"="3")) |>
  mutate(porosity = dplyr::recode(porosity, "71"="low", "81"="high")) |>
  rename("initialStrain"="strain") |>
  rename("initialStress"="stresskpa") |>
  mutate_at(c("Cyc_A", "Cyc_B", "Cyc_C", "Cyc_D", "Cyc_E", "Cyc_F", "Cyc_G", "Cyc_H", "Cyc_I", 
              "Cyc_J", "Cyc_K", "Cyc_L", "Cyc_M", "Cyc_N", "E1_MPA_Cyc_A", "E1_MPA_Cyc_B", 
              "E1_MPA_Cyc_C", "E1_MPA_Cyc_D", "E1_MPA_Cyc_E", "E1_MPA_Cyc_F", "E1_MPA_Cyc_G", 
              "E1_MPA_Cyc_H", "E1_MPA_Cyc_I", "E1_MPA_Cyc_J", "E1_MPA_Cyc_K", "E1_MPA_Cyc_L", 
              "E1_MPA_Cyc_M", "E1_MPA_Cyc_N", "delta_rad_Cyc_A", "delta_rad_Cyc_B", 
              "delta_rad_Cyc_C", "delta_rad_Cyc_D", "delta_rad_Cyc_E", "delta_rad_Cyc_F", 
              "delta_rad_Cyc_G", "delta_rad_Cyc_H", "delta_rad_Cyc_I", "delta_rad_Cyc_J", 
              "delta_rad_Cyc_K", "delta_rad_Cyc_L", "delta_rad_Cyc_M", "delta_rad_Cyc_N"), 
            as.numeric) |>
  mutate_at(c("E1_MPA_Cyc_A", "E1_MPA_Cyc_B", "E1_MPA_Cyc_C", "E1_MPA_Cyc_D", "E1_MPA_Cyc_E", 
              "E1_MPA_Cyc_F", "E1_MPA_Cyc_G", "E1_MPA_Cyc_H", "E1_MPA_Cyc_I", "E1_MPA_Cyc_J", 
              "E1_MPA_Cyc_K", "E1_MPA_Cyc_L", "E1_MPA_Cyc_M", "E1_MPA_Cyc_N"), abs)

#These lines add columns for Rate Of Change of stiffness and damping.
foamDataIntermediate <- foamDataIntermediate |>
  mutate(stiffness_ROC_A = NA) |>
  mutate(stiffness_ROC_B = (E1_MPA_Cyc_B - E1_MPA_Cyc_A)/(Cyc_B - Cyc_A))|>
  mutate(stiffness_ROC_C = (E1_MPA_Cyc_C - E1_MPA_Cyc_B)/(Cyc_C - Cyc_B))|>
  mutate(stiffness_ROC_D = (E1_MPA_Cyc_D - E1_MPA_Cyc_C)/(Cyc_D - Cyc_C))|>
  mutate(stiffness_ROC_E = (E1_MPA_Cyc_E - E1_MPA_Cyc_D)/(Cyc_E - Cyc_D))|>
  mutate(stiffness_ROC_F = (E1_MPA_Cyc_F - E1_MPA_Cyc_E)/(Cyc_F - Cyc_E))|>
  mutate(stiffness_ROC_G = (E1_MPA_Cyc_G - E1_MPA_Cyc_F)/(Cyc_G - Cyc_F))|>
  mutate(stiffness_ROC_H = (E1_MPA_Cyc_H - E1_MPA_Cyc_G)/(Cyc_H - Cyc_G))|>
  mutate(stiffness_ROC_I = (E1_MPA_Cyc_I - E1_MPA_Cyc_H)/(Cyc_I - Cyc_H))|>
  mutate(stiffness_ROC_J = (E1_MPA_Cyc_J - E1_MPA_Cyc_I)/(Cyc_J - Cyc_I))|>
  mutate(stiffness_ROC_K = (E1_MPA_Cyc_K - E1_MPA_Cyc_J)/(Cyc_K - Cyc_J))|>
  mutate(stiffness_ROC_L = (E1_MPA_Cyc_L - E1_MPA_Cyc_K)/(Cyc_L - Cyc_K))|>
  mutate(stiffness_ROC_M = (E1_MPA_Cyc_M - E1_MPA_Cyc_L)/(Cyc_M - Cyc_L))|>
  mutate(stiffness_ROC_N = (E1_MPA_Cyc_N - E1_MPA_Cyc_M)/(Cyc_N - Cyc_M))|>
  mutate(stiffness_frac_A = E1_MPA_Cyc_A / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_B = E1_MPA_Cyc_B / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_C = E1_MPA_Cyc_C / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_D = E1_MPA_Cyc_D / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_E = E1_MPA_Cyc_E / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_F = E1_MPA_Cyc_F / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_G = E1_MPA_Cyc_G / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_H = E1_MPA_Cyc_H / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_I = E1_MPA_Cyc_I / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_J = E1_MPA_Cyc_J / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_K = E1_MPA_Cyc_K / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_L = E1_MPA_Cyc_L / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_M = E1_MPA_Cyc_M / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_N = E1_MPA_Cyc_N / E1_MPA_Cyc_A) |>
  mutate(damping_frac_A = tan(delta_rad_Cyc_A) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_B = tan(delta_rad_Cyc_B) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_C = tan(delta_rad_Cyc_C) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_D = tan(delta_rad_Cyc_D) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_E = tan(delta_rad_Cyc_E) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_F = tan(delta_rad_Cyc_F) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_G = tan(delta_rad_Cyc_G) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_H = tan(delta_rad_Cyc_H) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_I = tan(delta_rad_Cyc_I) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_J = tan(delta_rad_Cyc_J) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_K = tan(delta_rad_Cyc_K) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_L = tan(delta_rad_Cyc_L) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_M = tan(delta_rad_Cyc_M) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_N = delta_rad_Cyc_N / delta_rad_Cyc_A) |>mutate(damping_ROC_A = NA) |>
  mutate(damping_ROC_B = (delta_rad_Cyc_B - delta_rad_Cyc_A)/(Cyc_B - Cyc_A))|>
  mutate(damping_ROC_C = (delta_rad_Cyc_C - delta_rad_Cyc_B)/(Cyc_C - Cyc_B))|>
  mutate(damping_ROC_D = (delta_rad_Cyc_D - delta_rad_Cyc_C)/(Cyc_D - Cyc_C))|>
  mutate(damping_ROC_E = (delta_rad_Cyc_E - delta_rad_Cyc_D)/(Cyc_E - Cyc_D))|>
  mutate(damping_ROC_F = (delta_rad_Cyc_F - delta_rad_Cyc_E)/(Cyc_F - Cyc_E))|>
  mutate(damping_ROC_G = (delta_rad_Cyc_G - delta_rad_Cyc_F)/(Cyc_G - Cyc_F))|>
  mutate(damping_ROC_H = (delta_rad_Cyc_H - delta_rad_Cyc_G)/(Cyc_H - Cyc_G))|>
  mutate(damping_ROC_I = (delta_rad_Cyc_I - delta_rad_Cyc_H)/(Cyc_I - Cyc_H))|>
  mutate(damping_ROC_J = (delta_rad_Cyc_J - delta_rad_Cyc_I)/(Cyc_J - Cyc_I))|>
  mutate(damping_ROC_K = (delta_rad_Cyc_K - delta_rad_Cyc_J)/(Cyc_K - Cyc_J))|>
  mutate(damping_ROC_L = (delta_rad_Cyc_L - delta_rad_Cyc_K)/(Cyc_L - Cyc_K))|>
  mutate(damping_ROC_M = (delta_rad_Cyc_M - delta_rad_Cyc_L)/(Cyc_M - Cyc_L))|>
  mutate(damping_ROC_N = (delta_rad_Cyc_N - delta_rad_Cyc_M)/(Cyc_N - Cyc_M))

# These lines lengthen the data by putting each measurement interval in its own data frame, 
#and row-binding them together.
foamDataA <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_A, E1_MPA_Cyc_A, 
                    stiffness_ROC_A, stiffness_frac_A, delta_rad_Cyc_A, damping_ROC_A, 
                    damping_frac_A, NVP_Cyc_A) |>
  rename("cycle"="Cyc_A") |>
  rename("stiffness"="E1_MPA_Cyc_A") |>
  rename("damping"="delta_rad_Cyc_A") |>
  rename("NVP"="NVP_Cyc_A") |>
  mutate("intervalCode"="A") |>
  rename("stiffnessROC"="stiffness_ROC_A") |>
  rename("stiffnessFrac"="stiffness_frac_A") |>
  rename("dampingROC"="damping_ROC_A") |>
  rename("dampingFrac"="damping_frac_A")
foamDataB <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_B, E1_MPA_Cyc_B, 
                    stiffness_ROC_B, stiffness_frac_B, delta_rad_Cyc_B, damping_ROC_B, 
                    damping_frac_B, NVP_Cyc_B) |>
  rename("cycle"="Cyc_B") |>
  rename("stiffness"="E1_MPA_Cyc_B") |>
  rename("damping"="delta_rad_Cyc_B") |>
  rename("NVP"="NVP_Cyc_B") |>
  mutate("intervalCode"="B") |>
  rename("stiffnessROC"="stiffness_ROC_B") |>
  rename("stiffnessFrac"="stiffness_frac_B") |>
  rename("dampingROC"="damping_ROC_B") |>
  rename("dampingFrac"="damping_frac_B")
foamDataC <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_C, E1_MPA_Cyc_C, 
                    stiffness_ROC_C, stiffness_frac_C, delta_rad_Cyc_C, damping_ROC_C, 
                    damping_frac_C, NVP_Cyc_C) |>
  rename("cycle"="Cyc_C") |>
  rename("stiffness"="E1_MPA_Cyc_C") |>
  rename("damping"="delta_rad_Cyc_C") |>
  rename("NVP"="NVP_Cyc_C") |>
  mutate("intervalCode"="C") |>
  rename("stiffnessROC"="stiffness_ROC_C") |>
  rename("stiffnessFrac"="stiffness_frac_C") |>
  rename("dampingROC"="damping_ROC_C") |>
  rename("dampingFrac"="damping_frac_C")
foamDataD <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_D, E1_MPA_Cyc_D, 
                    stiffness_ROC_D, stiffness_frac_D, delta_rad_Cyc_D, damping_ROC_D, 
                    damping_frac_D, NVP_Cyc_D) |>
  rename("cycle"="Cyc_D") |>
  rename("stiffness"="E1_MPA_Cyc_D") |>
  rename("damping"="delta_rad_Cyc_D") |>
  rename("NVP"="NVP_Cyc_D") |>
  mutate("intervalCode"="D") |>
  rename("stiffnessROC"="stiffness_ROC_D") |>
  rename("stiffnessFrac"="stiffness_frac_D") |>
  rename("dampingROC"="damping_ROC_D") |>
  rename("dampingFrac"="damping_frac_D")
foamDataE <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_E, E1_MPA_Cyc_E, 
                    stiffness_ROC_E, stiffness_frac_E, delta_rad_Cyc_E, damping_ROC_E, 
                    damping_frac_E, NVP_Cyc_E) |>
  rename("cycle"="Cyc_E") |>
  rename("stiffness"="E1_MPA_Cyc_E") |>
  rename("damping"="delta_rad_Cyc_E") |>
  rename("NVP"="NVP_Cyc_E") |>
  mutate("intervalCode"="E") |>
  rename("stiffnessROC"="stiffness_ROC_E") |>
  rename("stiffnessFrac"="stiffness_frac_E") |>
  rename("dampingROC"="damping_ROC_E") |>
  rename("dampingFrac"="damping_frac_E")
foamDataF <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_F, E1_MPA_Cyc_F, 
                    stiffness_ROC_F, stiffness_frac_F, delta_rad_Cyc_F, damping_ROC_F, 
                    damping_frac_F, NVP_Cyc_F) |>
  rename("cycle"="Cyc_F") |>
  rename("stiffness"="E1_MPA_Cyc_F") |>
  rename("damping"="delta_rad_Cyc_F") |>
  rename("NVP"="NVP_Cyc_F") |>
  mutate("intervalCode"="F") |>
  rename("stiffnessROC"="stiffness_ROC_F") |>
  rename("stiffnessFrac"="stiffness_frac_F") |>
  rename("dampingROC"="damping_ROC_F") |>
  rename("dampingFrac"="damping_frac_F")
foamDataG <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_G, E1_MPA_Cyc_G, 
                    stiffness_ROC_G, stiffness_frac_G, delta_rad_Cyc_G, damping_ROC_G, 
                    damping_frac_G, NVP_Cyc_G) |>
  rename("cycle"="Cyc_G") |>
  rename("stiffness"="E1_MPA_Cyc_G") |>
  rename("damping"="delta_rad_Cyc_G") |>
  rename("NVP"="NVP_Cyc_G") |>
  mutate("intervalCode"="G") |>
  rename("stiffnessROC"="stiffness_ROC_G") |>
  rename("stiffnessFrac"="stiffness_frac_G") |>
  rename("dampingROC"="damping_ROC_G") |>
  rename("dampingFrac"="damping_frac_G")
foamDataH <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_H, E1_MPA_Cyc_H, 
                    stiffness_ROC_H, stiffness_frac_H, delta_rad_Cyc_H, damping_ROC_H, 
                    damping_frac_H, NVP_Cyc_H) |>
  rename("cycle"="Cyc_H") |>
  rename("stiffness"="E1_MPA_Cyc_H") |>
  rename("damping"="delta_rad_Cyc_H") |>
  rename("NVP"="NVP_Cyc_H") |>
  mutate("intervalCode"="H") |>
  rename("stiffnessROC"="stiffness_ROC_H") |>
  rename("stiffnessFrac"="stiffness_frac_H") |>
  rename("dampingROC"="damping_ROC_H") |>
  rename("dampingFrac"="damping_frac_H")
foamDataI <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_I, E1_MPA_Cyc_I, 
                    stiffness_ROC_I, stiffness_frac_I, delta_rad_Cyc_I, damping_ROC_I, 
                    damping_frac_I, NVP_Cyc_I) |>
  rename("cycle"="Cyc_I") |>
  rename("stiffness"="E1_MPA_Cyc_I") |>
  rename("damping"="delta_rad_Cyc_I") |>
  rename("NVP"="NVP_Cyc_I") |>
  mutate("intervalCode"="I") |>
  rename("stiffnessROC"="stiffness_ROC_I") |>
  rename("stiffnessFrac"="stiffness_frac_I") |>
  rename("dampingROC"="damping_ROC_I") |>
  rename("dampingFrac"="damping_frac_I")
foamDataJ <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_J, E1_MPA_Cyc_J, 
                    stiffness_ROC_J, stiffness_frac_J, delta_rad_Cyc_J, damping_ROC_J, 
                    damping_frac_J, NVP_Cyc_J) |>
  rename("cycle"="Cyc_J") |>
  rename("stiffness"="E1_MPA_Cyc_J") |>
  rename("damping"="delta_rad_Cyc_J") |>
  rename("NVP"="NVP_Cyc_J") |>
  mutate("intervalCode"="J") |>
  rename("stiffnessROC"="stiffness_ROC_J") |>
  rename("stiffnessFrac"="stiffness_frac_J") |>
  rename("dampingROC"="damping_ROC_J") |>
  rename("dampingFrac"="damping_frac_J")
foamDataK <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_K, E1_MPA_Cyc_K, 
                    stiffness_ROC_K, stiffness_frac_K, delta_rad_Cyc_K, damping_ROC_K, 
                    damping_frac_K, NVP_Cyc_K) |>
  rename("cycle"="Cyc_K") |>
  rename("stiffness"="E1_MPA_Cyc_K") |>
  rename("damping"="delta_rad_Cyc_K") |>
  rename("NVP"="NVP_Cyc_K") |>
  mutate("intervalCode"="K") |>
  rename("stiffnessROC"="stiffness_ROC_K") |>
  rename("stiffnessFrac"="stiffness_frac_K") |>
  rename("dampingROC"="damping_ROC_K") |>
  rename("dampingFrac"="damping_frac_K")
foamDataL <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_L, E1_MPA_Cyc_L, 
                    stiffness_ROC_L, stiffness_frac_L, delta_rad_Cyc_L, damping_ROC_L, 
                    damping_frac_L, NVP_Cyc_L) |>
  rename("cycle"="Cyc_L") |>
  rename("stiffness"="E1_MPA_Cyc_L") |>
  rename("damping"="delta_rad_Cyc_L") |>
  rename("NVP"="NVP_Cyc_L") |>
  mutate("intervalCode"="L") |>
  rename("stiffnessROC"="stiffness_ROC_L") |>
  rename("stiffnessFrac"="stiffness_frac_L") |>
  rename("dampingROC"="damping_ROC_L") |>
  rename("dampingFrac"="damping_frac_L")
foamDataM <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_M, E1_MPA_Cyc_M, 
                    stiffness_ROC_M, stiffness_frac_M, delta_rad_Cyc_M, damping_ROC_M, 
                    damping_frac_M, NVP_Cyc_M) |>
  rename("cycle"="Cyc_M") |>
  rename("stiffness"="E1_MPA_Cyc_M") |>
  rename("damping"="delta_rad_Cyc_M") |>
  rename("NVP"="NVP_Cyc_M") |>
  mutate("intervalCode"="M") |>
  rename("stiffnessROC"="stiffness_ROC_M") |>
  rename("stiffnessFrac"="stiffness_frac_M") |>
  rename("dampingROC"="damping_ROC_M") |>
  rename("dampingFrac"="damping_frac_M")
foamDataN <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_N, E1_MPA_Cyc_N, 
                    stiffness_ROC_N, stiffness_frac_N, delta_rad_Cyc_N, damping_ROC_N, 
                    damping_frac_N, NVP_Cyc_N) |>
  rename("cycle"="Cyc_N") |>
  rename("stiffness"="E1_MPA_Cyc_N") |>
  rename("damping"="delta_rad_Cyc_N") |>
  rename("NVP"="NVP_Cyc_N") |>
  mutate("intervalCode"="N") |>
  rename("stiffnessROC"="stiffness_ROC_N") |>
  rename("stiffnessFrac"="stiffness_frac_N") |>
  rename("dampingROC"="damping_ROC_N") |>
  rename("dampingFrac"="damping_frac_N")

foamData <- rbind(foamDataA, foamDataB, foamDataC, foamDataD, foamDataE, foamDataF, foamDataG, 
                  foamDataH, foamDataI, foamDataJ, foamDataK, foamDataL, foamDataM, foamDataN)

# Filter rows with NA in the cycle, stiffness, and damping columns
foamData <- filter(foamData, !is.na(cycle))
